A Bidirected String Graph Model of Genome Assembly

### 2012-11-29 presentation/presentation.tex
Before update build graph (cont)

\begin{frame}{Genome assembly vs Shortest Common Substring}
	Genome assembly is superficially similar to the shortest common superstring
	(SCS) problem, which is to find the shortest string that contains all strings
	from a given set.
	\begin{center}
	\includegraphics[scale=0.6]{venn-crop.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Overlaps between reads}
	Consider two reads, $f$ and $g$, and their reverse-complements $f'$ and $g'$.
	The reads $f$ and $g$ are said to {\bf exactly overlap} by some length
	$\LengthVar$ iff at least one of the following is true:
	\begin{itemize}
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g'$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g'$.
	\end{itemize}
	An exact overlap $o$ is fully described by the tuple $(f, g, f_{beg},
	g_{beg}, len, rc)$.
	%there $f$ and $g$ are the reads of the overlap,
	%$f_{beg}$ and $g_{beg}$ and the starting position of the overlap on each
	%read, $len$ is the length of the overlap, and $rc$ indicates whether the
	%overlap is reverse-complement or not.
\end{frame}

@@

		\item Use the overlaps to construct a graph modeling how the reads could
		be put together
		\item Make each read be a vertex
		\item Make each edge be an overlap
		\begin{itemize}
			\item Two reads are connected with an edge iff they share an overlap.
		\end{itemize}
		\item Label each edge with an appropriate DNA sequence; this makes the
		graph a {\bf string graph}.
		\item Should the edges be directed or undirected?  \\Or $\dots$ something
		else?
	\end{itemize}
\end{frame}

@@

\begin{frame}{Walks in a bidirected graph}
	\begin{itemize}
		\item A {\bf walk in a bidirected graph} $G$ is a continuous sequence of
		edges in $G$ such that if any vertex $v$ is entered through a head
		inwards, it is left on a head outwards (unless it is the end of the
		path), and vice versa.
		\item A bidirected edge may be traversed in opposite directions during
		the same walk.
		\item The reverse of a walk in a bidirected graph is also a
		valid walk.
		\item $1 \to 2 \to 3 \to 4$ and $4 \to 3 \to 2 \to 1$ are both valid
		walks in the bidirected graph below.
	\end{itemize}
		\begin{center} {\small
		\begin{tikzpicture}[>=triangle 45]
			\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
			\node (1) at (0.00, 0) {1};
			\node (2) at (1.75, 0) {2};
			\node (3) at (3.50, 0) {3};
			\node (4) at (5.25, 0) {4};
			\draw[<->,very thick] (1) edge (2);
			\draw[>-<,very thick] (2) edge (3);
			\draw[<-<,very thick] (3) edge (4);
		\end{tikzpicture} } \end{center}
\end{frame}

\begin{frame}{Walking through the bidirected string graph}
	\begin{itemize}
		\item A walk through a bidirected string graph constructed from read
		overlaps models a way in which the reads can be assembled together
		consistently.

@@

		\begin{tikzpicture}[>=\ReadArrowType]
				\draw[<-,style=very thick] (0.0, 0.0) -- (4.0, 0.0);
				\draw[<-,style=very thick] (1.2, -0.3) -- (5.2, -0.3);
				\draw[->,style=very thick] (2.4, -0.6) -- (6.4, -0.6);
				\draw[<-,style=very thick] (3.6, -0.9) -- (7.6, -0.9);
				\draw (-0.3, 0.0) node {1.};
				\draw (-0.3, -0.3) node {2.};
				\draw (-0.3, -0.6) node {3.};
				\draw (-0.3, -0.9) node {4.};
		\end{tikzpicture}
		\caption{A bidirected graph and the implied assembly of the
		corresponding reads}
	\end{figure}
\end{frame}

>>>>

\begin{frame}{Genome assembly vs Shortest Common Superstring}
	Genome assembly is superficially similar to the shortest common superstring
	(SCS) problem, which is to find the shortest string that contains all strings
	from a given set.
	\begin{center}
	\includegraphics[scale=0.6]{venn-crop.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Overlaps between reads}
	Consider two reads, $f$ and $g$, and their reverse-complements $f'$ and $g'$.
	The reads $f$ and $g$ are said to {\bf exactly overlap} by some length
	$\LengthVar$ iff at least one of the following is true:
	\begin{itemize}
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f$ exactly match the first
		$\LengthVar$ bp of $g'$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g$.
		\item The last $\LengthVar$ bp of $f'$ exactly match the first
		$\LengthVar$ bp of $g'$.
	\end{itemize}
	An exact overlap $o$ is fully described by the tuple $(f, g, f_{begin},
	g_{begin}, L, rc)$.
	%there $f$ and $g$ are the reads of the overlap,
	%$f_{beg}$ and $g_{beg}$ and the starting position of the overlap on each
	%read, $len$ is the length of the overlap, and $rc$ indicates whether the
	%overlap is reverse-complement or not.
\end{frame}

@@

		\item Use the overlaps to construct a graph modeling how the reads could
		be put together
		\item Make each read be a vertex
		\item Make each edge be an overlap
		\begin{itemize}
			\item Two vertices are connected with an edge iff the
			corresponding reads share an overlap.
		\end{itemize}
		\item Label each edge with an appropriate DNA sequence; this makes the
		graph a {\bf string graph}.
		\item Should the edges be directed or undirected?  \\Or $\dots$ something
		else?
	\end{itemize}
\end{frame}

@@

\begin{frame}{Walks in a bidirected graph}
	\begin{itemize}
		\item A {\bf walk in a bidirected graph} $G$ is a continuous sequence of
		edges in $G$ such that if any vertex $v$ is entered through a head
		inwards, it is left on a head outwards (unless it is the end of the
		path), and vice versa.
		\item A bidirected edge may be traversed in both directions
		(possibly even during the same walk).
		\item The reverse of a walk in a bidirected graph is also a
		valid walk.
		\item $1 \to 2 \to 3 \to 4$ and $4 \to 3 \to 2 \to 1$ are both valid
		walks in the bidirected graph below.  (Check circled pairs of
		heads.)
	\end{itemize}
		\begin{center} {\small
		\begin{tikzpicture}[>=triangle 45]
			\tikzstyle{every node} = [circle,fill=blue!20,draw=black];
			\node (1) at (0.00, 0) {1};
			\node (2) at (2.00, 0) {2};
			\node (3) at (4.00, 0) {3};
			\node (4) at (6.00, 0) {4};
			\draw[<->,very thick] (1) edge (2);
			\draw[>-<,very thick] (2) edge (3);
			\draw[<-<,very thick] (3) edge (4);
			\draw[color=red,thick] (2.00, 0) circle (0.8);
			\draw[color=red,thick] (4.00, 0) circle (0.8);
		\end{tikzpicture} } \end{center}
\end{frame}

\begin{frame}{Walking through the bidirected string graph}
	\begin{itemize}
		\item A walk through a bidirected string graph constructed from
		reads and overlaps models a way in which the reads can be
		assembled together consistently.

@@

		\begin{tikzpicture}[>=\ReadArrowType]
				\draw[<-,style=very thick] (0.0, 0.0) -- (4.0, 0.0);
				\draw[->,style=very thick] (1.2, -0.3) -- (5.2, -0.3);
				\draw[<-,style=very thick] (2.4, -0.6) -- (6.4, -0.6);
				\draw[<-,style=very thick] (3.6, -0.9) -- (7.6, -0.9);
				\draw (-0.3, 0.0) node {1.};
				\draw (-0.3, -0.3) node {2.};
				\draw (-0.3, -0.6) node {3.};
				\draw (-0.3, -0.9) node {4.};
		\end{tikzpicture}
		\caption{A bidirected graph and the implied assembly of the
		corresponding reads}
	\end{figure}
\end{frame}
